---
title: "STA601Lab1"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose we want to estimate theta, the proportion of animals that can fly, from a random sample of animals.

```{r dataset}
library(cluster)
data(animals)
```



The indicators are coded to take on values 2, which represent the animal has the characteristic and 1 otherwise. We want to replace 2,1 into 1,0.

```{r transformation}
animals <- apply(animals, c(1, 2), function(x) {x - 1})
```

Since we're interested in theta, the proportion of animals that can fly, we should look at the number of animals in the sample that can fly.

```{r sum}
sum(animals[, "fly"])
```
So we have y=4 from a Binomial(n=20, theta) distribution.


Suppose we believe Suppose we believed:

big theta={0.01,…,0.99};
p(theta)=1/99 for each theta belongs to big theta.

This is a uniform prior on theta.

We can compute the posterior distribution using Bayes' rule.

```{r Bayes Rule}
y <- sum(animals[, "fly"])
n <- nrow(animals)

Theta.1 <- seq(0.01, 0.99, by = 0.01)
py.theta.1 <- dbinom(y, n, Theta.1)
ptheta.1 <- rep(1/length(Theta.1), length(Theta.1))
pytheta.1 <- py.theta.1*ptheta.1

ptheta.y.1 <- pytheta.1/(sum(pytheta.1))
```

Plot using Base R 

```{r Plot}
plot(Theta.1, ptheta.y.1, type = "l", 
     col = "blue",
     xlab = "theta",
     ylab = "p(theta|y)", 
     cex.lab = 1, cex.axis = 1)

lines(Theta.1, ptheta.1, col = "red", lty = 1)

abline(v = mean(animals[, "fly"]), lty = 2)

legend("topright", lty = c(1, 1, 2), 
       col = c("blue", "red", "black"),
       legend = c("p(theta|y)", "p(theta)", 
                  "y/n"))
```

Compute Posterior Mean and Variance
```{r Posterior Mean and Variance}
etheta.y.1 <- sum(ptheta.y.1*Theta.1)
etheta.y.1

vtheta.y.1 <- sum(ptheta.y.1*(Theta.1 - etheta.y.1)^2)
vtheta.y.1
```


Suppose we believe prior to be uniform distribution on interval (0,1)

This prior corresponds to the Beta(1, 1) distribution.

Because the beta distribution is conjugate for the binomial sampling model, this means that we know that the posterior distribution of theta will be a beta distribution, with parameters we have closed form solutions for posterior beta

```{r new param and plot}
atheta.y.2 <- 1 + y
btheta.y.2 <- 1 + n - y

Theta.2 <- seq(0.01, 0.99, by = 0.01)
ptheta.2 <- dbeta(Theta.2, 1, 1)
ptheta.y.2 <- dbeta(Theta.2, atheta.y.2, 
                    btheta.y.2)

plot(Theta.2, ptheta.y.2, type = "l", 
     col = "blue",
     xlab = "theta",
     ylab = "p(theta|y)", 
     cex.lab = 1, cex.axis = 1)

lines(Theta.2, ptheta.2, col = "red", lty = 1)
abline(v = mean(animals[, "fly"]), lty = 2)
legend("topright", lty = c(1, 1, 2), 
       col = c("blue", "red", "black"),
       legend = c("p(theta|y)", "p(theta)", 
                  "y/n"))
```

We can directly compute the posterior mean from what we know about the beta distribution.
```{r posterior mean and variance}

etheta.y.2 <- atheta.y.2/(atheta.y.2 + btheta.y.2)
etheta.y.2

vtheta.y.2 <- atheta.y.2*btheta.y.2/((atheta.y.2 + btheta.y.2)^2*(atheta.y.2 + btheta.y.2 + 1))
vtheta.y.2
```
Suppose we believed:

Big theta=(0,1);
Expectation of theta=0.43, i.e. about 2/5 of animals fly;
Variance of theta=0.03, i.e. we're pretty certain that theta around 0.43.

Again, we can directly compute the posterior mean, variance and other properties.
```{r prior we believe}
atheta.y.3 <- 3 + y
btheta.y.3 <- 4 + n - y

Theta.3 <- seq(0.01, 0.99, by = 0.01)
ptheta.3 <- dbeta(Theta.3, 3, 4)
ptheta.y.3 <- dbeta(Theta.3, atheta.y.3, 
                    btheta.y.3)

plot(Theta.3, ptheta.y.3, type = "l", 
     col = "blue",
     xlab = "theta",
     ylab = "p(theta|y)", 
     cex.lab = 1, cex.axis = 1)

lines(Theta.3, ptheta.3, col = "red", lty = 1)
abline(v = mean(animals[, "fly"]), lty = 2)
legend("topright", lty = c(1, 1, 2), 
       col = c("blue", "red", "black"),
       legend = c("p(theta|y)", "p(theta)", 
                  "y/n"))

etheta.y.3 <- atheta.y.3/(atheta.y.3 + btheta.y.3)
etheta.y.3

vtheta.y.3 <- atheta.y.3*btheta.y.3/((atheta.y.3 + btheta.y.3)^2*(atheta.y.3 + btheta.y.3 + 1))
vtheta.y.3
```